---
title: "Assignment-3"
author: "Sundeep Rachuri"
date: "2023-11-05"
---

```{r}
## Load required libraries
library(tidyverse)
library(reshape)
library(reshape2)
library(caret)
library(e1071)

## Load the data csv file
UniversalBank <- read.csv("C:/Users/snnra/Downloads/UniversalBank.csv")

head(UniversalBank)
tail(UniversalBank)
```

```{r}
## defining Column names
colnames(UniversalBank)

#Transforming data into categorical factors.

UniversalBank$`Personal Loan` = as.factor(UniversalBank$Personal.Loan)
UniversalBank$Online = as.factor(UniversalBank$Online)
UniversalBank$CreditCard = as.factor(UniversalBank$CreditCard)
```

```{r}
#Spliting the data into two the 60% of data in training set and 40% into validation set
set.seed(456)

# 60 % training data and 40% Validation data respectively
UniversalBank.train.data <- sample(row.names(UniversalBank), 0.6*dim(UniversalBank)[1])
UniversalBank.valid.data <- setdiff(row.names(UniversalBank), UniversalBank.train.data)

# assigning training data and validation data to a dataframe from dataset respectively
UniversalBank.train <- UniversalBank[UniversalBank.train.data, ]
UniversalBank.valid <- UniversalBank[UniversalBank.valid.data, ]

# Duplicating the data frame UB.train
train <- UniversalBank[UniversalBank.train.data, ]
# Duplicating the data frame UB.valid
valid = UniversalBank[UniversalBank.train.data,]
```

## A. Create a pivot table for the training data with Online as a column variable, CC as a row variable, and Loan as a secondary row variable. The values inside the table should convey the count. In R use functions melt() and cast(), or function table().

```{r}
## melt data from data
melted.data = melt(train,id=c("CreditCard","Personal.Loan"),variable= "Online")

# Casting the melted data
cast.data <- dcast(melted.data, CreditCard + Personal.Loan ~ value, fun.aggregate = length)

# Casting column number 14: Personal loan, ID, and credit card, respectively
cast.data[,c(1,2,3,14)]
```

#B. Consider the task of classifying a customer who owns a bank credit card and is actively using online banking services. Looking at the pivot table, what is the probability that this customer will accept the loan offer? [This is the probability of loan acceptance (Loan = 1) conditional on having a bank credit card (CC = 1) and being an active user of online banking services (Online = 1)].

```{r}
#According to the pivot table, the value for Loan is 89, and the value for CC is 
Universal.Bank.Loan.CC1 <- 89/3000

# which is 29 %.
Universal.Bank.Loan.CC1
```
## C. Create two separate pivot tables for the training data. One will have Loan (rows) as a function of Online (columns) and the other will have Loan (rows) as a function of CC.

```{r}
 ## reshaping the train data frame into a long format, with "Personal.Loan" as an identifier and "Online" as a variable to be melted
melt1.data = melt(train,id=c("Personal.Loan"),variable = "Online")

 ## reshaping the train data frame into a long format, with "CreditCard" as an identifier and "Online" as a variable to be melted
melt2.data = melt(train,id=c("CreditCard"),variable = "Online")
```

```{r}
# Casting Personal loan and online values
# Assuming both Personal.Loan and Online are binary (0 or 1)
cast1.data =dcast(melt1.data,`Personal.Loan`~Online) 

# Casting Personal loan and online values
cast2.data=dcast(melt2.data,CreditCard~Online)
```

```{r}
#shows the number of personal loans in reference to online
UniversalBank.Loanonline=cast1.data[,c(1,13)]
UniversalBank.LoanCC = cast2.data[,c(1,14)]
UniversalBank.Loanonline
```
```{r}

# shows the number of credit cards in reference to Online.
UniversalBank.LoanCC

```
## D. Compute the following quantities [P (A | B) means “the probability of A given B”]:
#1.P (CC = 1 | Loan = 1) (the proportion of credit card holders among the loan acceptors) #2.P(Online=1|Loan=1) 
#3.P (Loan = 1) (the proportion of loan acceptors)
#4.P(CC=1|Loan=0)
#5.P(Online=1|Loan=0)
#6.P(Loan=0)

```{r}
# creating a pivot table with the columns 14 and 10 representing personal loan
table(train[,c(14,10)])
```
```{r}
# Creating a pivot table for column 13 and 10 which is online and personal loan

table(train[,c(13,10)])
```

```{r}
 # Personal loan pivot table There are 2725 and 275 from training, respectively
table(train[,c(10)])
```
```{r}
## We can obtain the CC= 1 and Loan = 1 values by referring to the above p
UniversalBank.CCUniversalBank.Loan1 = 89/(89+200)
UniversalBank.CCUniversalBank.Loan1
```
2. P(Online=1|Loan=1)

```{r}
##We can get the online = 1 and loan = 1 values from the pivot table aboveUB.ONUB.Loan1
UniversalBank.ONUniversalBank.Loan1 =177/(177+112)
```


3. P(Loan = 1)

```{r}
 # By referring the above pivot table we can get the Loan = 1
UniversalBank.Loan1 =289/(289+2711)
UniversalBank.Loan1
```


4. P(CC=1|Loan=0)

```{r}

#Using the pivot table above, we can obtain the CC = 1 and Loan = 0 values
UniversalBank.CCLoan.01= 794/(794+1917) 
UniversalBank.CCLoan.01
```

5. P(Online=1|Loan=0)

```{r}
# We can get the online = 1 and loan = 0 values from the pivot table above.
UniversalBank.ON1.L0= 1665/(1665+1046)
UniversalBank.ON1.L0
```

6. P(Loan=0)

```{r}
# We can obtain the Loan = 0 values by the pivot table above.
UniversalBank.Loan0= 2711/(2711+289)
UniversalBank.Loan0
```
##E. Use the quantities computed above to compute the naive Ba1 probability P(Loan = 1 | CC = 1, Online = 1).

```{r}
# Given probabilities
P_CC_given_Loan1 <- 0.096
P_Online_given_Loan1 <- 0.833
P_Loan1 <- 0.0125

# Calculate Naive Bayes probability P(Loan = 1 | CC = 1, Online = 1)
#Universal_Bank_Naivebayes <- (P_CC_given_Loan1 * P_Online_given_Loan1 * P_Loan1) / (P_CC * P_Online)

UniversalBank_Naivebayes <- (P_CC_given_Loan1)*(P_Online_given_Loan1)*(P_Loan1)
UniversalBank_Naivebayes

```

## F. Compare this value with the one obtained from the pivot table in (b). Which is a more accurate estimate?

```{r}
# Naive Bayes Probability (from calculation in E)
naive_bayes_probability <- 0.0009996

# Pivot Table Probability (from part B)
pivot_table_probability <- 0.02966667

# Compare the probabilities and print a message indicating which is more accurate
if (naive_bayes_probability > pivot_table_probability) {
  message("Naive Bayes Probability is more accurate: ", naive_bayes_probability)
} else if (naive_bayes_probability < pivot_table_probability) {
  message("Pivot Table Probability is more accurate: ", pivot_table_probability)
} else {
  message("Both probabilities are the same: ", naive_bayes_probability)
}

```
##Based on the comparison, the pivot table probability (0.02966667) is considered more accurate compared to the Naive Bayes probability


## G. Which of the entries in this table are needed for computing P(Loan = 1 | CC = 1, Online = 1)? Run naive Bayes on the data. Examine the model output on training data, and find the entry that corresponds to P(Loan = 1 | CC = 1, Online = 1). Compare this to the number you obtained in (E).

```{r}

names(UniversalBank)

names(UniversalBank.train)

# Select the relevant columns for training
UniversalBank.train <- UniversalBank[, c("CreditCard", "Online", "Personal Loan")]

# Rename the columns to remove spaces
colnames(UniversalBank.train) <- c("CreditCard", "Online", "PersonalLoan")

# Convert 'CreditCard' and 'Online' to factors with appropriate levels
UniversalBank.train$CreditCard <- factor(UniversalBank.train$CreditCard, levels = c(0, 1), labels = c("No", "Yes"))
UniversalBank.train$Online <- factor(UniversalBank.train$Online, levels = c(0, 1), labels = c("No", "Yes"))
```


# Print the probability
print("Probability of Loan = 1 given CC = 1 and Online = 1:")
print(probability_loan_1)

```


